### 2025-05-23 docs/article/Sections/state-of-the-art.tex
Chapter 2: add Spring Webflux subsection

>>>>

Spring WebFlux is a reactive web framework built on a non-blocking,
event-driven architecture that enables efficient handling of a large number of
concurrent connections with minimal resource usage. Unlike traditional
servlet-based stacks that rely on a thread-per-request model, WebFlux operates
on reactive runtimes like Netty, where a small, fixed-size event loop manages
all I/O events. Incoming HTTP requests are routed to handler functions or
annotated controllers, which return reactive types such as \texttt{Mono<T>} and
\texttt{Flux<T>}—Publisher implementations defined by the Reactive Streams
specification. A \texttt{Mono} represents a single asynchronous value (or
none), while a \texttt{Flux} represents a stream of zero or more values. These
abstractions allow the entire request lifecycle—from controller invocation to
response rendering—to be orchestrated as a chain of non-blocking, asynchronous
operations. I/O-bound tasks such as database queries or external API calls do
not occupy threads during execution, enabling the framework to scale gracefully
under high load and maintain responsiveness even in resource-constrained
environments.

Progressive Server-Side Rendering (PSSR) is natively supported in WebFlux via
\texttt{Flux}-based controllers. When a controller returns a
\texttt{Flux<String>} or another stream of HTML fragments, WebFlux can begin
streaming content to the client as soon as the first elements are emitted. This
allows browsers to incrementally render parts of the response without waiting
for the entire payload, significantly reducing Time to First Byte (TTFB) and
improving user-perceived performance. By integrating PSSR directly into the
reactive processing pipeline, WebFlux facilitates responsive and resilient web
applications well-suited for modern microservices architectures.

An illustrative example of PSSR in WebFlux is shown below:

\begin{lstlisting}[language=Java]
@GetMapping(value = "/pssr", produces = MediaType.TEXT_HTML_VALUE)
public Flux<String> renderChunks() {
    return Flux.range(1, 10)
               .delayElements(Duration.ofMillis(500))
               .map(i -> "<p>Chunk " + i + "</p>\n");
}
\end{lstlisting}

In this example, the method emits one HTML chunk every 500 milliseconds. As
each chunk is generated, it is immediately written to the HTTP response,
allowing the client to begin rendering content even as more data is still being
produced. Unlike \texttt{StreamingResponseBody} in Spring MVC, which is
constrained by servlet buffer thresholds and blocking semantics, WebFlux
ensures that each emitted item is pushed to the client as soon as the network
is ready, without occupying a dedicated thread. Moreover, WebFlux integrates
seamlessly with reactive data sources such as R2DBC (Reactive Relational
Database Connectivity) and reactive NoSQL drivers, making it possible to build
end-to-end non-blocking applications. HTML builders and DSLs used in
conjunction with WebFlux must be designed to accommodate reactive streams,
ensuring that view generation does not violate the non-blocking contract by
performing blocking I/O or awaiting asynchronous operations imperatively.
